---
title: "seabirds analysis"
output: html_notebook
---


```{r}
library(tidyverse)
seabirds <- read_csv(here("clean_data/sea_birds_clean.csv"))
seabirds

```

the Wandering albatross sensu lato is the bird that had the most individual sightings with 11293 sightings across the dataset

```{r}
seabirds %>% group_by(common_name) %>% 
summarize(individual_sightings = n()) %>% 
arrange(desc(individual_sightings))

```

bird with the highest total count is the Short-tailed shearwater with a total count of 982553	

```{r}
seabirds %>% group_by(common_name) %>% 
summarize(total_count = sum(count, na.rm = TRUE)) %>%
arrange(desc(total_count))
```
the bird that had the highest total count above a latitude of -30 was the Wedge-tailed shearwater with the highest total count of 1360 out of records above -30 degrees
```{r}
seabirds %>%
  filter(lat >= -30) %>%
  group_by(common_name) %>% 
  summarize(total_count = sum(count, na.rm = TRUE)) %>%
  arrange(desc(total_count))
```
How many different types of birds were only ever seen in groups of 1?
Atlantic yellow-nosed mollymawk
Caspian tern
Chatham Island mollymawk	
Dark Procellaria (unidentified)		
Stejneger's petrel		
Yellow-eyed penguin
were all seen exclusivly without a group
```{r}
seabirds %>% group_by(common_name) %>%
tally(ocacc == "N") %>%
rename(times_seen_in_group = n) %>%
arrange(times_seen_in_group)
```
How many penguins were seen?:
penguins were seen on 70 different occasions
the total number of penguins counted was 158
```{r}
all_penguins = filter(seabirds, str_detect(common_name, "[pP]enguin"))
nrow(all_penguins)
all_penguins %>%
summarize(total_count_of_penguins = sum(count, na.rm = TRUE))
```

