---
title: "seabirds analysis"
output: html_notebook
---


```{r}
library(tidyverse)
seabirds <- read_csv(here("clean_data/sea_birds_clean.csv"))
seabirds

```

all following questions assume that each birds common name refers only to one species of bird.

the Wandering albatross sensu lato is the bird that had the most individual sightings with 11293 sightings across the dataset

```{r}
seabirds %>% group_by(common_name) %>% 
summarize(individual_sightings = n()) %>% 
arrange(desc(individual_sightings))

```

bird with the highest total count is the Short-tailed shearwater with a total count of 982553	

```{r}
seabirds %>% group_by(common_name) %>% 
summarize(total_count = sum(count, na.rm = TRUE)) %>%
arrange(desc(total_count))
```
the bird that had the highest total count above a latitude of -30 was the Wedge-tailed shearwater with the highest total count of 1360 out of records above -30 degrees
```{r}
seabirds %>%
  filter(lat >= -30) %>%
  group_by(common_name) %>% 
  summarize(total_count = sum(count, na.rm = TRUE)) %>%
  arrange(desc(total_count))
```
How many different types of birds were only ever seen in groups of 1?
the following birds were all seen exclusivly without a group 
```{r}
seabirds %>% group_by(common_name) %>%
tally(ocacc == "Y" | ocacc == "U") %>%
rename(times_seen_in_group = n) %>%
arrange(times_seen_in_group) %>%
filter(times_seen_in_group == 0) %>%
select(common_name)
```
How many penguins were seen?:
penguins were seen on 70 different occasions
the total number of penguins counted was 158
```{r}
all_penguins = filter(seabirds, str_detect(common_name, "[pP]enguin"))
nrow(all_penguins)
all_penguins %>%
summarize(total_count_of_penguins = sum(count, na.rm = TRUE))
```

